Experience Report: Distributed Computing for the Common Man

### 2010-03-25 doc/icfp-2010/main.tex
petit corrections, but it introduced an orphan :( [KK]

\authorinfo{Jean-Christophe Filliâtre}
           {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
            INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
           {filliatr@lri.fr}
\authorinfo{Kalyan Krishnamani}
           {INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
           {kalyan@lri.fr}

@@

The application domain of such a distributed computing library is
manyfold.  The most striking application, as also relevant to our
research endeavours, is in validating thousands of verification
conditions using automated theorem provers. Having noted this, it
should be apparent that the library finds use in efficient execution
of any large-scale expensive computation that is parallelizable.

% what it is and what it is not
% \begin{itemize}
% \item it is not a library which does parallelization
% \item it is a library for reliable distributed execution of
%   parallelizable computation 
% \end{itemize}
% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
% programming, ironically) ;  however, there are differences:
% \begin{itemize}
% \item polymorphic, higher-order API
% \item more generic (cores/network)
% \item does not focus on association lists
% \item we don't have data locality (future work)
% \end{itemize}
% what is the target audience/applications: 
% for instance, use if automatic provers on thousands of verification
% conditions, on computing infrastructures which can be
% \begin{itemize}
% \item a single machine with multiple cores, possibly remote
% \item several machines, small or large in number, over a network
% \end{itemize}

\paragraph{Distributed Computing.}
A typical distributed computing library provides the following (we
borrow some terminology from Google's Mapreduce~\cite{mapreduce}):
\begin{itemize}
\item A notion of a \emph{task} which denotes the atomic computation to be
  performed in a distributed manner;
\item A set of machines or programs called \emph{workers} that perform
  the tasks, producing results;
\item A single program called a \emph{master} that is in charge
  of distributing the tasks among the workers and managing results
  produced by the workers.
\end{itemize}
In addition to the above, distributed computing environments also
implement mechanisms for fault-tolerance, efficient storage and
distribution of tasks. This is required to handle network failures
that may occur, as well as to optimize the usage of machines in the
network. Another concern of importance is the transmission of messages
over the network. This requires efficient serialization (or
marshalling) of data that can be reliably reconstructed when
transmitted over different computing environments.  It is desirable to
maintain architecture independence while transmitting marshalled data,
as machines in a distributed computing environment often run on
different hardware architectures and make use of varying software
platforms. For example, endianness may be different across machines on
the network.

@@

\section{API}\label{sec:API}

The main function in our API follows the idea sketched in the
introduction. It has the following signature:
\begin{ocaml}
  val compute : 
    worker:('a -> 'b) -> 
    master:('a * 'c -> 'b -> ('a * 'c) list) -> 
    ('a * 'c) list -> unit
\end{ocaml}
Tasks are pairs, of type \of{'a * 'c}, where the first component is
passed to the worker and the second component is local to the master.
The function \ocaml{worker} should be pure and is executed in parallel
with other \ocaml{worker}s. The function \ocaml{master}, on the
contrary, can be impure and is never executed in parallel with other
\ocaml{master}s. The function \ocaml{master} always processes
information locally, typically storing it to internal data structures.
The next section will describe the usage of this generic interface to
perform traditional map and fold operations.

@@

In this section, we show the potential of our library with two 
computationally-extensive examples.

@@

The first column shows the value of $N$.
Each computation is divided into a certain number of tasks,
which is shown in second column.
Then the last three columns show execution timings in seconds for the
three implementations. The figures within brackets show the speedup
ratio w.r.t. sequential implementation.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    N & \#tasks  & sequential& cores                 & network 
    \\\hline\hline
    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
    \\\hline
       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
    \\\hline
    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
    \\\hline
       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
    \\\hline
    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
    \\\hline
       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
    \\\hline
    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
    \\\hline
       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
    \\\hline
  \end{tabular}
\end{center}
From the table above, it is clear that the \of{Cores} and \of{Network}
implementations
provides a significant speedup. As evident from the last row,
the speedup ratio is almost 8, which is also the number of
cores we use. 
It is also evident from the last column that \of{Network}
implementation performs significantly better when the computation time
dominates the total execution time.
The two extreme cases correspond to second and last rows: in the
second row, the communication time dominates and is in fact more than
91\%\ of the total execution time; on the other hand, for the last row
communication time amounts to just 4.6\%\ of the total execution time.
As expected, the network implementation is only beneficial when the
computation time for each individual task is significant, which is the
case in realistic examples.

@@

The master program proceeds by reading the 80 file names, turning them
into tasks by multiplying them by the number of provers (240 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 10 minutes.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{cannot-decide} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
The figure below shows the total time in minutes spent by each prover
for each possible outcomes.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & cannot & timeout & failure
    \\\hline\hline
    Z3       &   4.20 &   0.00 &    160  &  20.0    
    \\\hline
    Alt-ergo &   3.60 &   0.57 &    190  &   8.2     
    \\\hline
    Simplify &   0.02 &   0.00 &     40  &   1.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 7 hours if provers were executed
sequentially. However, using our library and our 3-machines
infrastructure, it completes in 1 hour and 5 minutes, giving us a
speedup of 6.6$\times$.

@@

Closest to the approach in this paper is Yohann Padioleau's mapreduce
implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
top of OcamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing.  On one hand, we have less flexibility
w.r.t. deployment of the user program; on the other hand, we provide a
more generic API together with fault tolerance.
There are other distributed computing libraries on top of which one
could implement the library discussed in this paper. \JoCaml\
is one of them~\cite{jocaml}. However, \JoCaml\ does not allow
arbitrary closures to be passed between machines, which would limit
the genericity of the API. Additionally, \JoCaml\ does not provide
fault tolerance, which is essential in a distributed setting.
% TODO check with Louis

>>>>

\authorinfo{Jean-Christophe Filli\^{a}tre}
           {CNRS, LRI, Univ Paris-Sud, Orsay F-91405\\
            INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
           {filliatr@lri.fr}
\authorinfo{Kalyan Krishnamani}
           {INRIA Saclay - Île-de-France, ProVal, Orsay, F-91893}
           {kalyan@lri.fr}

@@

The application domain of such a distributed computing library is
manyfold.  The most striking application, as also relevant to our
research endeavours, is in validating thousands of verification
conditions using automated theorem provers. Having noted this, it
should be apparent that the library finds use in efficient execution
of any large-scale expensive computation that is parallelizable.
% what it is and what it is not
% \begin{itemize}
% \item it is not a library which does parallelization
% \item it is a library for reliable distributed execution of
%   parallelizable computation 
% \end{itemize}
% inspired by Google's MapReduce~\cite{mapreduce} (itself inspired by functional
% programming, ironically) ;  however, there are differences:
% \begin{itemize}
% \item polymorphic, higher-order API
% \item more generic (cores/network)
% \item does not focus on association lists
% \item we don't have data locality (future work)
% \end{itemize}
% what is the target audience/applications: 
% for instance, use if automatic provers on thousands of verification
% conditions, on computing infrastructures which can be
% \begin{itemize}
% \item a single machine with multiple cores, possibly remote
% \item several machines, small or large in number, over a network
% \end{itemize}
\paragraph{Distributed Computing.}
A typical distributed computing library provides the following (we
borrow some terminology from Google's Mapreduce~\cite{mapreduce}):
\begin{itemize}
\item A notion of a \emph{task} which denotes the atomic computation to be
  performed in a distributed manner;
\item A set of machines or programs called \emph{workers} that perform
  the tasks, producing results;
\item A single program called a \emph{master} that is in charge
  of distributing the tasks among the workers and managing results
  produced by the workers.
\end{itemize}
In addition to the above, distributed computing environments also
implement mechanisms for fault-tolerance, efficient storage and
distribution of tasks. This is required to handle network failures
that may occur, as well as to optimize the usage of machines in the
network. Another concern of importance is the transmission of messages
over the network. This requires efficient serialization (or
marshalling) of data that can be reliably reconstructed when
transmitted over different computing environments.  It is desirable to
maintain architecture independence while transmitting marshalled data,
as machines in a distributed computing environment often run on
different hardware architectures and make use of varying software
platforms. For example, endianness may be different across machines on
the network.

@@

\section{API}\label{sec:API}
In this section, we describe the uniqueness of approach, namely the
generic API which we shall extend to a plethora of more specialized
APIs in later sections. The main function in our API follows the idea
sketched in the introduction. It has the following signature:
\begin{ocaml}
  val compute : 
    worker:('a -> 'b) -> 
    master:('a * 'c -> 'b -> ('a * 'c) list) -> 
    ('a * 'c) list -> unit
\end{ocaml}
Tasks are pairs, of type \of{'a * 'c}, where the first component is
passed to the worker and the second component is local to the master.
The function \ocaml{worker} should be pure and is executed in parallel
with other \ocaml{worker}s. The function \ocaml{master}, on the
contrary, can be impure and is never executed in parallel with other
\ocaml{master}s. The function \ocaml{master} always processes
information locally, typically storing it to internal data structures.
The next section will describe the usage of this generic interface to
perform traditional map and fold operations.

@@

In this section, we show the potential of our library using two 
computationally-extensive examples.

@@

The first column shows the value of $N$.
Each computation is divided into a certain number of tasks,
which is shown in second column.
Then the last three columns show execution timings in seconds for the
three implementations. The figures within brackets show the speedup
ratio w.r.t. sequential implementation.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    N & \#tasks  & sequential& cores                 & network 
    \\\hline\hline
    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
    \\\hline
       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
    \\\hline
    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
    \\\hline
       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
    \\\hline
    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
    \\\hline
       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
    \\\hline
    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
    \\\hline
       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
    \\\hline
  \end{tabular}
\end{center}
From the table above, it is clear that the \of{Cores} and \of{Network}
implementations provide a significant speedup. As evident from the
last row, the speedup ratio is almost 8, which is also the number of
cores we use.  It is also evident from the last column that
\of{Network} implementation performs significantly better when the
computation time dominates the total execution time.  The two extreme
cases correspond to second and last rows: in the second row, the
communication time dominates and is in fact more than 91\%\ of the
total execution time; on the other hand, for the last row
communication time amounts to just 4.6\%\ of the total execution time.
As expected, the network implementation is only beneficial when the
computation time for each individual task is significant, which is the
case in realistic examples.

@@

The master program proceeds by reading the 80 file names, turning them
into tasks by multiplying them by the number of provers (240 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 10 minutes.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{cannot-decide} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
The figure below shows the total time in minutes spent by each prover
for each possible outcomes.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & cannot & timeout & failure
    \\\hline\hline
    Z3       &   4.20 &   0.00 &    160  &  20.0    
    \\\hline
    Alt-ergo &   3.60 &   0.57 &    190  &   8.2     
    \\\hline
    Simplify &   0.02 &   0.00 &     40  &   1.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 7 hours if provers were executed
sequentially. However, using our library and our 3-machine
infrastructure, it completes in 1 hour and 5 minutes, giving us a
speedup of 6.6$\times$.

@@

Closest to the approach in this paper is Yohann Padioleau's mapreduce
implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
top of OcamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing.  On one hand, we have less flexibility
w.r.t. deployment of the user program; on the other hand, we provide a
more generic API together with fault tolerance.  There are other
distributed computing libraries on top of which one could implement
the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
them. However, \JoCaml\ does not allow arbitrary closures to be passed
between machines, which would limit the genericity of the
API. Additionally, \JoCaml\ does not provide fault tolerance, which is
indispensible in a distributed setting.
% TODO check with Louis
